\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant floral and reproductive traits determine interactions with floral visitors

### 2021-08-13 Draft/Manuscript.Rmd
editys

**Plant strategies.** The phylogenetically informed principal component analysis (pPCA) captured by the first two and three axes 51.8% and 70.97% of trait variation, respectively (Fig. 1 and Supplementary Fig. S5) and had a phylogenetic correlation ($\lambda$) of 0.76. The first principal component (PC1) represented 26.72% of the trait variation and indicated a trade-off between flower number and flower size. We refer to this axis as the ‘flower number - flower size trade-off as already described in previous studies [@sargent2007; @kettle2011]. Hence, one end of the spectrum comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings > |0.5|; Supplementary Table S3) but style length also contributed moderately on PC1 (loading = -0.33). The second principal component (PC2) represented 25.05% of the trait variation and indicated a trade-off between low and high pollinator dependence. We refer to this axis as the ‘pollinator dependence trade-off’. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Supplementary Table S3). We found that high pollinator dependence was associated with larger and higher numbers of flowers, greater plant height and longer styles. In contrast, species with high levels of autonomous selfing tended to have fewer and smaller flowers, had shorter styles and were shorter in height. Further, PC3 explained a considerable amount of trait variability (19.17%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Supplementary Table S3). Thus, because style length was correlated with all traits on PC3 and was the main driver of variation, we refer to this axis as the ‘style length trade-off’. Further, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) was positively associated with flower size, style length and ovule number (PC1, 23.40%); and pollen quantity (pollen grains per flower) was positively correlated with flower number and plant height and negatively associated with autonomous selfing (PC2, 21.67%; Supplementary Fig. S6). This pPCA explained similar variance with the first two principal components (45.07%) and similar associations of traits despite some variability in the loadings (Supplementary Table S4).

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,fig.pos = "t",fig.cap= "**Fig. 1 |** Life history strategies for 1,236 plant species from 28 plant-pollinator networks studies across the first two axes of trait variation from a phylogenetically informed principal component analysis (pPCA). The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) in the 2-d plane. The length of the arrows indicate the weight of the variables on each principal component and the labelled icons at their end represent the extreme form of the trait continuum. The dashed lines show the opposed direction of trait variation and the non-labelled icons at their end illustrate the opposing extreme of the continuum."}

@@

We found that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Supplementary Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, p < 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , p = 0.59; PC2:  Sum of squares = 0.83, F-value = 1.43, p = 0.23) showed lack of statistical association. In addition, we found statistical differences (Tukey test) between the different levels of categorical traits in the trait space (Supplementary Fig. S7). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatibile were statistically differentiated from species with partial or full self compatibility; Supplementary Fig. S7 A and B). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Supplementary Fig. S7 C and D). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Supplementary Fig. S7 E and F). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Supplementary Fig. S7 G and H). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Supplementary Fig. S7 I and J). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Supplementary Fig. S7 K and L).

@@

**Phylogenetic signal of traits.** We found a strong phylogenetic signal (P < 0.01) in all quantitative traits (Supplementary Table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda = 1$), pollen grains per flower ($\lambda = 1$) and plant height ($\lambda = 0.96$), followed by flower length ($\lambda = 0.75$), flower width ($\lambda = 0.73$), number of flowers per plant ($\lambda = 0.69$) and nectar concentration ($\lambda = 0.65$). The traits that showed a moderate phylogenetic signal were inflorescence width ($\lambda = 0.57$), style length ($\lambda = 0.49$) and autonomous selfing ($\lambda = 0.34$). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all traits ($\lambda = 0.14$).

**Visitation patterns.** The main axes of trait variation explained little of the overall visitation rates (conditional $R2 = 0.31$; marginal $R2 = 0.06$) but showed relevant trends when we explored the interaction with the different floral visitors guilds (Fig. 3). All floral visitors guilds visited plant species with higher pollinator dependence more frequently (PC2; Fig. 3b and Fig. 3e). In addition, all guilds other than Syrphids and Lepidoptera (i.e., all Hymenoptera, non-Syrphidae-Diptera and Coleoptera) showed greater visitation rates on species with small numerous flowers (PC1; Fig. 3a and Fig. 3d). On the style length trade-off (PC3), Anthophila-Hymenoptera, Lepidoptera and Non-Anthophila-Hymenoptera showed greater visitation rates on plant species with larger styles and higher levels of selfing; while Syrphidae, Non-Syrphidae-Diptera and Coleoptera showed higher visitation rates on species with shorter styles and lower selfing (Fig. 3c and Fig. 3f). Furthermore, when running an additional model that separates the most represented families of Anthophila-Hymenoptera (bees; $marginal R2 = 0.30$; $conditional R2 = 0.03$) showed that the family Apidae was the main driver of the observed patterns (Supplementary Fig. S8).

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= "**Fig. 2 | Fitted posterior estimates of the number of visits made by the different floral visitors guilds in relation to the three main axes of trait variation (PC1, PC2 and PC3)**. Bees (a, b and c) and the rest of guilds (d, e and f) were plotted separately for visualization purposes. In addition, we trimmed the plotting area that was over the 95th percentile to improve visualisation. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off."}

@@

**Plant species functional roles.** The three main axes of trait variation poorly explain the variance of the different plant species level metrics (Supplementary Fig. S9; interaction frequency ~ PCs, $conditional R2 = 0.11$, $marginal R2 = 0.02$; normalized degree ~ PCs, $conditional R2 = 0.24$, $marginal R2 = 0.02$; and, specialization ~ PCs, $conditional R2 = 0.37$, $marginal R2 = 0.03$). Overall, the most notable trends were found on PC1 and PC3 for interaction frequency and specialization. On the flower number - flower size trade-off (PC1), interaction frequency was higher for plant species with more flowers but was lower for plant species with larger flowers. On PC1, specialization showed the opposite trend. On the style length trade-off (PC3), interaction frequency was lower for plants with shorter styles and lower selfing and higher for species with long styles and higher levels of selfing. Again, specialization showed the opposite trend to interaction frequency.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.pos = "!H",fig.cap= "**Fig. 4 | Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits**. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

#Set correct colnames for plotting
colnames(v_df) <- c("Visits", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule number","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen grains per flower")

colnames(nd_df) <- c("Normalized degree", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule number","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen grains per flower")

colnames(d_df) <- c("Specialization", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule number","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen grains per flower")

@@

rpart.plot(tree1, box.palette="GnOr", main="(a) Interaction frequency", cex.main=1.5,tweak = 1.25,type = 2) 
rpart.plot(tree2, box.palette="GnOr", main="(b) Normalized degree", cex.main=1.5,tweak = 1.25,type = 2)
rpart.plot(tree3, box.palette="GnOr", main="(c) Specialization", cex.main=1.5,tweak = 1.25,type = 2)

@@

Over 50% of plant trait variation was captured by the flower number - flower size and pollinator dependence trade-offs. Variation in these two axes of trait variation appears to be associated with the commonly described ‘fast-slow continuum’ in plant [@salguero2016] and animal [@healy2019] life history strategies as indicated by the different floral and reproductive traits associated with plant height, life form and lifespan. The “slow” part of the continuum (i.e., tall trees and shrubs) had plant species with many flowers, few ovules, higher pollinator dependence, greater occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, species that employed the “fast” strategy (i.e., short herbs), had fewer flowers, more ovules, greater occurrence of self compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the positive association between outcrossing rate and floral display, in accordance with @goodwillie2010. This trend was further confirmed when looking at the specific correlation between floral display and autonomous selfing for our dataset (see Supplementary Fig. S10).

Despite the low predictive power of the main axes of trait variation on overall visitation patterns, we found commonalities among and within floral visitors guilds across these axes that indicate the influence of life history strategies on floral visitation rate. For example, all guilds showed an increasing trend of visitation on plant species with higher pollinator dependence, which were also associated with larger floral displays and greater pollen quantity (Fig. 1 and Supplementary Fig. S6). This trend is consistent with previous plant-pollinator studies that show higher visitation rates on species that make greater reproductive investment [@hegland2005; @lazaro2013; @kaiser2014]. In regard to the flower number - flower size and style length trade-offs, different guilds showed contrasting visitation rates, which could be associated with different pollination syndromes at a macroecological scale. For instance, the guilds of bees and syrphids were clearly associated with opposing life-strategies on PC1 and PC3 (Fig. 3) indicating a possible niche partitioning among these two guilds [@palmer2003; @phillips2020]. However, these plant-pollinator associations do not account for some of the traits (i.e., pollen and nectar quantity) that had greater relevance on the species network roles (Fig. 4) because of insufficient data availability and lack of pollinator efficiency measurements for more complete descriptions of the pollination syndromes [@dellinger2020]. In any case, it is worth noting that this general pattern emerges in field-measured plant-pollinator networks, where other local factors such as species relative abundances, will surely explain part of the observed variability [@vazquez2007; @encinas2012; @bartomeus2016].

>>>>

**Plant strategies.** The phylogenetically informed principal component analysis (pPCA) captured by the first two and three axes 51.8% and 70.97% of trait variation, respectively (Fig. 1 and Supplementary Fig. S5) and had a phylogenetic correlation ($\lambda$) of 0.76. The first principal component (PC1) represented 26.72% of the trait variation and indicated a trade-off between flower number and flower size. We refer to this axis as the ‘flower number - flower size trade-off’ as already described in previous studies [@sargent2007; @kettle2011]. Hence, one end of the spectrum comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings > |0.5|; Supplementary Table S3) but style length also contributed moderately on PC1 (loading = -0.33). The second principal component (PC2) represented 25.05% of the trait variation and indicated a trade-off between low and high pollinator dependence. We refer to this axis as the ‘pollinator dependence trade-off’. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Supplementary Table S3). We found that high pollinator dependence was associated with larger and a higher number of flowers, greater plant height and longer styles. In contrast, species with high levels of autonomous selfing tended to have fewer and smaller flowers, had shorter styles and were shorter in height. Further, PC3 explained a considerable amount of trait variability (19.17%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Supplementary Table S3). Thus, because style length was correlated with all traits on PC3 and was the main driver of trait variation, we refer to this axis as the ‘style length trade-off’. Further, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) was positively associated with flower size, style length and ovule number (PC1, 23.40%); and pollen quantity (pollen grains per flower) was positively correlated with flower number and plant height and negatively associated with autonomous selfing (PC2, 21.67%; Supplementary Fig. S6). This pPCA explained similar variance with the first two principal components (45.07%) and similar associations of traits despite some variability in the loadings (Supplementary Table S4).

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,fig.pos = "t",fig.cap= "**Fig. 1 | Plant life-history strategies.** Phylogenetically informed principal component analysis (pPCA) of 1,236 plant species from 28 plant-pollinator networks studies. The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) across the two main axes of trait variation. The length of the arrows indicate the weight of the variables on each principal component and the labelled icons at their end represent the extreme form of the trait continuum. The dashed lines show the opposed direction of trait variation and the non-labelled icons at their end illustrate the opposing extreme of the continuum."}

@@

We found that most categorical traits were statistically associated with the first two axes of trait variation (Fig. 2 and Supplementary Table S2). Flower symmetry, which was only associated with PC2 (Sum of squares = 8.51, F-value = 14.72, _P_ < 0.01 ), and nectar provision, which was independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 , _P_ = 0.59; PC2:  Sum of squares = 0.83, F-value = 1.43, _P_ = 0.23) showed lack of statistical association. In addition, we found with Tukey test statistical differences between the different levels of categorical traits in the trait space (Supplementary Fig. S7). Regarding self compatibility, we found larger differences on PC2 (i.e., species with unisexual flowers that were self incompatibile were statistically differentiated from species with partial or full self compatibility; Supplementary Fig. S7a and Fig. S7b). Life forms differed statistically across both axes of trait variation and followed a gradient of larger life forms (trees and shrubs) with higher pollinator dependence to smaller ones (herbs) with lower pollinator dependence (Supplementary Fig. S7c and Fig. S7d). Consequently, lifespan also followed this gradient but perennial and short lived species only differed statistically on PC2 (Supplementary Fig. S7e and Fig. S7f). Species with unisexual flowers (monoecious and dioecious) were clustered on both extremes of the first two principal components and had the highest pollinator dependence and highest number of flowers (Supplementary Fig. S7g and Fig. S7h). Moreover, we found that the campanulate and capitulum flower shapes were differentiated from tube, papilionaceous, open and brush shapes in the trait space. The former morphologies had larger flowers and greater pollinator dependence, while the latter had higher flower number and greater autonomous selfing (Supplementary Fig. S7i and Fig. S7j). Regarding flower symmetry, zygomorphic flowers were associated with lower levels of pollinator dependence, whereas actinomorphic flowers had higher levels of pollinator dependence (Supplementary Fig. S7k and Fig. S7l).

@@

**Phylogenetic signal of traits.** We found a strong phylogenetic signal (*P* < 0.01) in all quantitative traits (Supplementary Table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda = 1$), pollen grains per flower ($\lambda = 1$) and plant height ($\lambda = 0.96$), followed by flower length ($\lambda = 0.75$), flower width ($\lambda = 0.73$), number of flowers per plant ($\lambda = 0.69$) and nectar concentration ($\lambda = 0.65$). The traits that showed a moderate phylogenetic signal were inflorescence width ($\lambda = 0.57$), style length ($\lambda = 0.49$) and autonomous selfing ($\lambda = 0.34$). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all traits ($\lambda = 0.14$).

**Visitation patterns.** The main axes of trait variation explained little of the overall visitation rates ($conditional R2 = 0.31$; $marginal R2 = 0.06$) but showed relevant trends when we explored the interaction with the different floral visitors guilds (Fig. 3). All floral visitors guilds visited plant species with higher pollinator dependence more frequently (PC2; Fig. 3b and Fig. 3e). In addition, all guilds other than Syrphids and Lepidoptera (i.e., all Hymenoptera, non-Syrphidae-Diptera and Coleoptera) showed greater visitation rates on species with small numerous flowers (PC1; Fig. 3a and Fig. 3d). On the style length trade-off (PC3), Anthophila-Hymenoptera, Lepidoptera and Non-Anthophila-Hymenoptera showed greater visitation rates on plant species with larger styles and higher levels of selfing; while Syrphidae, Non-Syrphidae-Diptera and Coleoptera showed higher visitation rates on species with shorter styles and lower selfing (Fig. 3c and Fig. 3f). Furthermore, when running an additional model that separates the most represented families of Anthophila-Hymenoptera (bees; $marginal R2 = 0.30$; $conditional R2 = 0.03$) showed that the family Apidae was the main driver of the observed patterns (Supplementary Fig. S8).

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.pos = "!H",fig.cap= "**Fig. 3 | Visitation rates across the three main axes of trait variation.** Fitted posterior estimates of the number of visits made by the different floral visitors guilds in relation to PC1, PC2 and PC3. Bees (a, b and c) and the rest of guilds (d, e and f) were plotted separately for visualization purposes. In addition, we trimmed the plotting area that was over the 95th percentile to improve visualisation. PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off."}

@@

**Plant species functional roles.** The variance of the different plant species level metrics was poorly explained by the three main axes of trait variation (Supplementary Fig. S9; interaction frequency ~ PCs, $conditional R2 = 0.11$, $marginal R2 = 0.02$; normalized degree ~ PCs, $conditional R2 = 0.24$, $marginal R2 = 0.02$; and, specialization ~ PCs, $conditional R2 = 0.37$, $marginal R2 = 0.03$). Overall, the most notable trends were found on PC1 and PC3 for interaction frequency and specialization. On the flower number - flower size trade-off (PC1), interaction frequency was higher for plant species with more flowers but was lower for plant species with larger flowers. On PC1, specialization showed the opposite trend. On the style length trade-off (PC3), interaction frequency was lower for plants with shorter styles and lower autonomous selfing and higher for species with longer styles and higher autonomous selfing. Again, specialization showed the opposite trend to interaction frequency.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.pos = "!H",fig.cap= "**Fig. 4 | Trait relevance on the plant species’ functional roles.** Regression tree analysis of interaction frequency (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

#Set correct colnames for plotting
colnames(v_df) <- c("Visits", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule n.","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen g. per flower")

colnames(nd_df) <- c("Normalized degree", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule n.","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen g. per flower")

colnames(d_df) <- c("Specialization", "Aut. selfing", "Flowers per plant", "Flower width", "Style length (mm)",
                      "Ovule n.","Plant height (m)",  "Microlitres of nectar",  "Nectar concentration (%)","Pollen g. per flower")

@@

rpart.plot(tree1, box.palette="GnOr", main="(a) Interaction frequency", cex.main=1.45,tweak = 1.25,type = 2) 
rpart.plot(tree2, box.palette="GnOr", main="(b) Normalized degree", cex.main=1.45,tweak = 1.25,type = 2)
rpart.plot(tree3, box.palette="GnOr", main="(c) Specialization", cex.main=1.45,tweak = 1.25,type = 2)

@@

Over 50% of plant trait variation was captured by the flower number - flower size and pollinator dependence trade-offs. Variation in these two axes of trait variation appears to be associated with the commonly described ‘fast-slow continuum’ in plant [@salguero2016] and animal [@healy2019] life-history strategies as indicated by the different floral and reproductive traits associated with plant height, life form and lifespan. The ‘slow’ part of the continuum (i.e., tall trees and shrubs) had plant species with many flowers, few ovules, higher pollinator dependence, greater occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, species that employed the ‘fast’ strategy (i.e., short herbs), had fewer flowers, more ovules, greater occurrence of self compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the **previously described** positive association between outcrossing rate and floral display [@goodwillie2010]. This trend was further confirmed when looking at the specific correlation between floral display and autonomous selfing for our dataset (see Supplementary Fig. S10).

Despite the low predictive power of the main axes of trait variation on overall visitation patterns, we found commonalities among and within floral visitors guilds across these axes that indicate the influence of life-history strategies on floral visitation rate. For example, all guilds showed an increasing trend of visitation on plant species with higher pollinator dependence, which were also associated with larger floral displays and greater pollen quantity (Fig. 1 and Supplementary Fig. S6). This trend is consistent with previous plant-pollinator studies that show higher visitation rates on species that make greater reproductive investment [@hegland2005; @lazaro2013; @kaiser2014]. In regard to the flower number - flower size and style length trade-offs, different guilds showed contrasting visitation rates, which could be associated with different pollination syndromes at a macroecological scale. For instance, the guilds of bees and syrphids were clearly associated with opposing life-strategies on PC1 and PC3 (Fig. 3) indicating a possible niche partitioning among these two guilds [@palmer2003; @phillips2020]. However, these plant-pollinator associations do not account for some of the traits (i.e., pollen and nectar quantity) that had greater relevance on the species network roles (Fig. 4) because of insufficient data availability and lack of pollinator efficiency measurements for more complete descriptions of the pollination syndromes [@dellinger2020]. In any case, it is worth noting that this general pattern emerges in field-measured plant-pollinator networks, where other local factors such as species relative abundances, will surely explain part of the observed variability [@vazquez2007; @encinas2012; @bartomeus2016].
